---
title: "Reproducible Research"
author: "NicolasGarcesR"
date: "7/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(ggplot2)
library(dplyr)
```

## Loading and preprocessing the data

```{r, echo=TRUE}

base<- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character", "integer"))
base$date <- ymd(base$date)
head(base)
```

## What is mean total number of steps taken per day?

- Calculate the total number of steps taken per day
```{r pressure, echo=T}
steps <- base %>%
  group_by(date) %>%
  summarize(steps = sum(steps,na.rm = T))
steps
```

- Make a histogram of the total number of steps taken each day

```{r}
hist(steps$steps, col= "blue", main = "Histogram of steps", xlab = "Total steps" )
```

- Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(steps$steps, na.rm = T)
median(steps$steps, na.rm = T)
```

## What is the average daily activity pattern?

- Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
interval <- base %>%
  group_by(interval) %>% 
  summarize(steps = mean(steps, na.rm = T))

plot(interval$interval,interval$steps, type = "l", col="blue", main = "time series of the 5-minute interval average steps", xlab = "interval", ylab = "steps")

```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval[max(interval$steps),]
```

## Imputing missing values

- Calculate and report the total number of missing values in the dataset

```{r}
sum(is.na(base$steps))

```

- Devise a strategy for filling in all of the missing values in the dataset.
```{r}
na.steps<- base$steps                          
steps.mean<- mean(na.steps, na.rm = TRUE)        
na.steps[is.na(na.steps)] <- steps.mean
base$steps <- as.numeric(na.steps)
```

- Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
steps1 <- base%>%
  group_by(date) %>%
  summarize(steps = sum(steps))

```

- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
hist(steps1$steps, col= "red", main = "Histogram of steps", xlab = "Total steps" )
mean(steps1$steps)
median(steps1$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

- Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
base<- mutate(base, weektype = ifelse(weekdays(base$date)== "Saturday" |weekdays(base$date)== "Sunday", "weekend", "weekday"))

head(base)
```

- Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
interval1 <- base %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
plot<- ggplot(interval1, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
plot
```

